---
title: "Blight prediction model"
author: "Mladen Cucak"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction


```{r}
# TODO Check outbreaks who have the same nearest station. 
# TODO Caclualte yearly weather and risk summaries
# TODO Try to figure out if there is a correlation with patterns of the weather and the climate. 
# This could be interesting when determining the population growth - the increase of the risk over time
# Probably need to make histograms for all stations to check if all of the data is ok. 

```

# Methods
## Model implementation


## Sporulation

The sporulation is suppressed by light (De Weille, 1963)
Add the effect of wind and rain - check norwegian work. 

Major influence on sporulation: 
temperature and relative humidity
Light cycle
wind
Function for sporulation in relation with relative humidity and temperature
Observe sporulation during the night  and then look into conditions for infection

Multiplication factor based on the sporulation from the night before
Sporulation stops at noon or when the conditions are not fulfilled any more 

## Infection 
Starts after 10 hours of sporulation and is assessed untill following day at 12 or untill conditions are not met any more. 
 



## Survival
The spores of the pathogen are ....by the sunlight (De Weille, 1963, Mizubuti et al, 2000, Skelsey 2017). 

##Validation
Do we want to test the sensitivity or do the calibration? Calibrating the model based on the data is misleading practice - because of the agronomy data has could have hiden variables that are not acounted for and could skew th biological principles behind the model. 

Model was evaluated using ...2.... sets of historical late blight outbreak data. The outbreak data spanned a 12-year period (2003-2014) and consisted of the date and coordinates of .....number of outbreaks... late blight outbreaks from across Northern Ireland. The outbreak data are collected and blight samples are send each year by blight scouts as part of the Agriculture and Horticulture Development Board (AHDB) Potatoes “Fight Against Blight” campaign (https://potatoes.ahdb.org.uk/). The coordinates of locations were found using geocoding function of ggmap package (Reference, ). 


Additional validation will be implemented based on the difference in number of treatments recomended by the model and 


#Weather forecast
http://www.agrometeorology.org/files-folder/repository/gamp_chapt5.pdf 
Measure certainy of decision with regards of steps in time. 
Run simulations with individual variables from the forecasted data (Kim, 2018)
Include spraying conditions evaluation - Detlefsen. 
CART/SLD, was first developed by Gleason et al. (23) and then adjusted by Kim et al. (30). CART/SLD has inputs of wind speed, dew point depression (the difference between air temperature and dew point temperature), air temperature, and RH.

Quantity of information difference between estimated and measured solar radiation. 


# Dissucion
Necessary to establish upper temperature limit due to possible effect of the climate change   
In the course of developent of hte present model a number of simplifications has been made to find an optimum balance between mimicking the life cycle of the pathogen and the scarcity of the weather network.  

Sustaining the activity may hinge upon continued dialogue and collaboration between the four core groups – Meteorologist, Research Officer, Extension Officer, Farmer





